---
title: "Lecture 1: Esimating Probability using Simulations"
output: 
  learnr::tutorial:
    theme: "sandstone"
    css: css/styles.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library("learnr")
knitr::opts_chunk$set(echo = TRUE)
```

## Basic R for the course

#### With R, you can 

+ Compute simple arithmetic operations:

```{r rcanadd}
2 / 5 + (5 - 4) * 1 
```

+  Evaluate logical statements:

```{r rcanlogic}
5 > 2 && 3 != 5 && 2 == 2
```

+  Simulate and plot data:

```{r rcanplot}
plot(rnorm(25))
```

In this class, we will use R to demonstrate probability concepts via simulation.

### Variables and data types

#### You can store values to variables such as 

+ numbers:

```{r variables-num}
my_number <- 5
my_number
```

+ characters:

```{r variables-char}
my_characters <- "this is a single sentence."
my_characters
```

+ logicals:

```{r variables-logic}
my_logical <- TRUE
my_logical
```

+ sequences of a single data type called vectors:

```{r vectors}
my_vector <- c(2, 3, 7)
my_vector
```

#### To extract elements from a vector, you can use 

+ the positional indices inside `[` starting with 1:

```{r vector-extract}
my_vector[1]
my_vector[c(1, 3)]
```

+ a logical vector of the same length to extract elements at `TRUE` positions:

```{r vector-extract-logical}
my_vector[c(FALSE, FALSE, TRUE)]
my_vector[c(T, F, T)]
```

#### You can create empty vectors of length `n` with:

```{r emptyvectors}
n <- 5
numeric(n) # numbers
character(n) # characters
logical(n) # logicals
```


### Practice

Create a vector that consists of day of your birthday and your first name. Print the vector and comment on its data type.

```{r vectors-ex, exercise=TRUE}
birth_name <- # create the vector
```

```{r vectors-ex-solution}
birth_name <- c(1, "Michael")
birth_name
```

### Functions

In R, a function is in the following form:

```
<function name>(<arg1>, <arg2>, ...)
```

+  You can store the result of a function with `result <- function_x(...)`

```{r function-sum, exercise=TRUE}
sum_of_237 <- sum(c(2, 3, 7))
sum_of_237
```

+  You can look at the help page of each function with `?<function name>`.

```{r function-ex, exercise=TRUE, echo=TRUE}
?sample

```

+  There are also a plenty of resources and examples online for most R functions.

### Loops and vectorized calls

To loop through each element in a vector, you may use a "for loop" in R:

```{r forloop}
x <- c(10, 20, 30, 40, 50)
y <- c(5, 10, 15, 20, 25)
z <- numeric(5)
for (i in c(1, 2, 3, 4, 5)) {
  z[i] <- x[i] + y[i]
}
z
```

Many R functions automatically broadcast the operation to each element resulting in a vector of same size.

```{r vectorized}
x + y
```


## Simulating Relative Frequency

Recall, one interpretation of _probability_ is long-term relative frequency. Suppose you repeat a random experiment a large number of time, say $N$, and observe $n$ number of event $A$ The relative frequency 
$$\frac{n}{N}$$ 
is approximately the probability $P(A)$.

### Example: Rolling a Fair Die 


Recall the single die example with
$$A=\text{You roll an even number}$$
and
$$B=\text{You roll a number less than 3}$$

We can simulate a single roll with.

```{r singleroll}
sample(x = 1:6, size = 1)

```

> `1:6` is a shortcut for `c(1, 2, 3, 4, 5, 6)`.

To simulate $N$ times, we can use a for loop.

```{r diceloop, exercise=TRUE, exercise.lines=8}
N <- 25
rolls <- numeric(N)
for(i in 1:N) {
  # simulate a single single roll and save to `rolls` at index `i`
}
rolls
```

Alternatively, we can ask `sample` to simulate $N$ times at once.

```{r dicevectorized, exercise=TRUE, exercise.lines=5}
N <- 25
rolls <- sample(1:6, N) # correct this line appropriately
rolls
```

```{r dicevectorized-solution}
N <- 25
rolls <- sample(1:6, N, replace = TRUE) 
# replace = TRUE allows drawing values from 1:6 multiple times
rolls
```

### Computing $P(A)$

In order to compute $P(A)$, we will count the relative frequency of simulated values that are even. We also want to use a large $N$. In the code chunk below, `N` is the number of simulated rolls, 1,000, and vector `rolls` contains 1,000 simulated rolls of a fair die. Use the vector to estimate $P(A)$ and print it.

```{r calcpa-setup, echo=FALSE}
set.seed(237)
N <- 1000
rolls <- sample(1:6, N, replace = TRUE)
```

```{r calcpa, exercise=TRUE, exercise.lines=10}
rolls

```

```{r calcpa-solution}
sum(rolls %% 2 == 0) / N
```

```{r calcpa-hint}
rolls %% 2 # results in a vector containing the remainder after dividing each roll by 2
```

### Practice: Computing $P(B)$

Estimate $P(B)$ based on 100,000 simulations without using a for loop.

```{r calcpb, exercise=TRUE, exercise.lines=10}


```

```{r calcpb-solution}
N <- 100000
rolls <- sample(1:6, N, replace = TRUE)
sum(rolls < 3) / N
```

<!-- ## Multiple Experiments -->

<!-- Suppose you roll a fair die twice. Let $E$ be the event the sum of the two rolls is greater than or equal to 8. -->

<!-- ### Compute by Simulation -->

<!-- ```{r calcpe, exercise=TRUE, exercise.lines=10} -->
<!-- N <- 10^6 -->
<!-- # replicate outputs result of `n` repetitions of  -->
<!-- # the operations described in `expr` -->
<!-- E <- replicate(  -->
<!--   n = N, expr = { -->
<!--     # roll a die twice -->
<!--     roll <- sample(1:6, size = 2, replace = TRUE)  -->
<!--     # output whether the sum is greater than or equal to 8 -->
<!--     sum(roll)  -->
<!--   }) -->
<!-- sum(E >= 8)/N -->
<!-- ``` -->

