---
title: "Lecture 1: Esimating Probability using Simulations"
output: 
  learnr::tutorial:
    css: css/styles.css
runtime: shiny_prerendered
---

<!-- 
Before running the document, make sure `learnr` package is available. 
If missing, you can install the package with `install.packages('learnr').

http://jupyter.utoronto.ca/hub/user-redirect/git-pull?repo=https%3A%2F%2Fgithub.com%2Fsta237%2Frlesson01&urlpath=shiny%2Frlesson01%2Frlesson01.Rmd&branch=main
-->

```{r setup, include=FALSE}
renv::activate(".")
renv::restore(".")
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Simulating Relative Frequency

Recall, one interpretation of _probability_ is long-term relative frequency. Suppose you repeat a random experiment a large number of time, say $N$, and observe $n$ number of event $A$ The relative frequency 
$$\frac{n}{N}$$ 
is approximately the probability $P(A)$.

### Example: Rolling a Fair Die 


Recall the single die example with
$$A=\text{You role an even number}$$
and
$$B=\text{You role a number less than 3}$$

We can simulate a single role with.

```{r singlerole, exercise=TRUE, exercise.lines=3}

sample(x = 1:6, size = 1)

```

To simulate $N$ times, we can use a for loop.

```{r echo=TRUE}
N <- 10
draw <- numeric(N)
for(i in 1:N) {
  draw[i] <- sample(1:6, 1)
}
print(draw)
```

Alternatively, we can ask `sample` to simulate $N$ times at once.

```{r echo=TRUE}
N <- 15
# replace must be set to TRUE to allow simulating a single value multiple times
draw <- sample(1:6, N, replace = TRUE) 
print(draw)
```

### Computing $P(A)$

In order to compute $P(A)$, we will count the relative frequency of simulated values that are even. We also want to use a large $N$.

```{r calcpa, exercise=TRUE, exercise.lines=10}
N <- 10^6 # simulate one million times
draw <- sample(1:6, N, replace = TRUE)


```


### Practice: Computing $P(B)$

Estimate $P(B)$ based on 100,000 simulations.

```{r calcpb, exercise=TRUE, exercise.lines=10}


```


## Multiple Experiments

Suppose you roll a fair die twice. Let $C$ be the event the sum of the two rolls is greater than or equal to 8.

### Compute by Simulation

```{r calcpc, exercise=TRUE, exercise.lines=10}
N <- 10^6
# replicate outputs result of `n` repetitions of 
# the operations described in `expr`
C <- replicate( 
  n = N, expr = {
    # roll a die twice
    roll <- sample(1:6, size = 2, replace = TRUE) 
    # output whether the sum is greater than or equal to 8
    sum(roll) >=8
  })
sum(C)/N
```

### Compute by Counting

Compute $P(C)$ by counting. Does the theoretical values match the simulated estimate?

_Hint: When the first roll is 1, the sum cannot be greater than or equal to 8. When the first roll is 2, the sum is greater than or equal to 8 when the second roll is 6. How many outcomes satisfy the event $C$ when the first roll is 3?_

